---
title: "The first project task"
output:
  html_document:
    df_print: paged
date: "2022-11-21"
---
```{r library}
library(ggplot2)
library(tidyr)
library(GGally)

```




## 1 User function
```{r function}
tablel_merge <- function(path_to_data) {
  names <- list.files(path=path_to_data, full.names=TRUE)
  tables <- lapply(names, function (x) read.csv(file=x, sep=','))
  Reduce(function(x,y) merge(x,y, all = TRUE), tables)}
```

## 2 A brief EDA

Downloading data:

```{r final_table}
work_table <- tablel_merge('/Users/asabukreeva/Downloads/Data')
```

EDA

```{r summary}
summary(work_table)

```
*Table check *
Working with NA in numeric columns
```{r NA mean}
work_table$Length <- as.numeric(work_table$Length) 
work_table$Length[is.na(work_table$Length)]<-mean(work_table$Length, na.rm = T)
work_table$Diameter[is.na(work_table$Diameter)]<-mean(work_table$Diameter, na.rm = T)
work_table$Height[is.na(work_table$Height)]<-mean(work_table$Height, na.rm = T)
work_table$Whole_weight[is.na(work_table$Whole_weight)]<-mean(work_table$Whole_weight, na.rm = T)
work_table$Shucked_weight[is.na(work_table$Shucked_weight)]<-mean(work_table$Shucked_weight, na.rm = T)
work_table$Shell_weight[is.na(work_table$Shell_weight)]<-mean(work_table$Shell_weight, na.rm = T)
work_table <- work_table %>% drop_na()

summary(work_table) 
```
Now I have the missing values in all numeric columns replaced by the column average.
Next step is correcting the incorrect values in the columns.
```{r correct}
work_table$Rings[(work_table$Rings) == 'nine']<- 9
work_table$Sex..1...male..2...female..3...uvenil.[(work_table$Sex..1...male..2...female..3...uvenil.) == 'male']<- 1
work_table$Sex..1...male..2...female..3...uvenil.[(work_table$Sex..1...male..2...female..3...uvenil.) == 'one']<- 1
work_table$Sex..1...male..2...female..3...uvenil.[(work_table$Sex..1...male..2...female..3...uvenil.) == 'three']<- 3
```
```{r remove NA}
work_table <- work_table %>% drop_na()
```
I decided to change names and values of sex column (original name is long and really annoying)
```{r sex}
work_table$Sex..1...male..2...female..3...uvenil.[(work_table$Sex..1...male..2...female..3...uvenil.) == 1]<- 'male'
work_table$Sex..1...male..2...female..3...uvenil.[(work_table$Sex..1...male..2...female..3...uvenil.) == 2]<- 'female'
work_table$Sex..1...male..2...female..3...uvenil.[(work_table$Sex..1...male..2...female..3...uvenil.) == 3]<- 'uvenil'
names(work_table)[names(work_table) == 'Sex..1...male..2...female..3...uvenil.'] <- 'Sex'
names(work_table)
```
I'm not sure that it is correct, but I decided to change Rings collumn as numeric. 
```{r Diameter_num}
work_table$Rings <- as.numeric(work_table$Rings)
```
Fine, it looks like the table is now clean and I can build boxplots to check for outliers.


*Checking Emissions*
```{r Length}
boxplot(work_table$Length,
        ylab = "Length")
```
```{r Diameter}
boxplot(work_table$Diameter,
        ylab = "Diameter")
```
```{r Height}
boxplot(work_table$Height,
        ylab = "Height")
```
```{r Whole_weight}
boxplot(work_table$Whole_weight,
        ylab = "Whole_weight")
```
```{r Shucked_weight}
boxplot(work_table$Shucked_weight,
        ylab = "Shucked_weight")
```
```{r Viscera_weight}
boxplot(work_table$Viscera_weight,
        ylab = "Viscera_weight")
```
```{r Shell_weight}
boxplot(work_table$Shell_weight,
        ylab = "Shell_weight")
```
```{r Rings}
boxplot(work_table$Rings,
        ylab = "Rings")
```





Oh, the outliers are everywhere. I try to get rid of them using quantile restrictions. 

```{r Length_outlier}
quartiles <- quantile(work_table$Length, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(work_table$Length)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

table_no_outlier <- subset(work_table, work_table$Length > Lower & work_table$Length < Upper)
boxplot(table_no_outlier$Length,
        ylab = "Length")
```
```{r Diameter_outlier}
quartiles <- quantile(table_no_outlier$Diameter, probs=c(.25, .75), na.rm = FALSE) 
IQR <- IQR(table_no_outlier$Diameter)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

table_no_outlier <- subset(table_no_outlier, table_no_outlier$Diameter> Lower & table_no_outlier$Diameter < Upper)
boxplot(table_no_outlier$Diameter,
        ylab = "Diameter")
```
```{r Height_outlier}
quartiles <- quantile(table_no_outlier$Height, probs=c(.25, .75), na.rm = FALSE) 
IQR <- IQR(table_no_outlier$Height)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

table_no_outlier <- subset(table_no_outlier, table_no_outlier$Height> Lower & table_no_outlier$Height < Upper)
boxplot(table_no_outlier$Height,
ylab = "Height")
```
```{r Whole_weight_outlier}
quartiles <- quantile(table_no_outlier$Whole_weight, probs=c(.25, .75), na.rm = FALSE) 
IQR <- IQR(table_no_outlier$Whole_weight)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

table_no_outlier <- subset(table_no_outlier, table_no_outlier$Whole_weight> Lower & table_no_outlier$Whole_weight < Upper)
boxplot(table_no_outlier$Whole_weight,
        ylab = "Whole weight") 
```
```{r Shucked_weight_outlier}
quartiles <- quantile(table_no_outlier$Shucked_weight, probs=c(.25, .75), na.rm = FALSE) 
IQR <- IQR(table_no_outlier$Shucked_weight)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

table_no_outlier <- subset(table_no_outlier, table_no_outlier$Shucked_weight> Lower & table_no_outlier$Shucked_weight < Upper)
boxplot(table_no_outlier$Shucked_weight,
        ylab = "Shucked weight")
```
```{r Viscera_weight_outlier}
quartiles <- quantile(table_no_outlier$Viscera_weight, probs=c(.25, .75), na.rm = FALSE) 
IQR <- IQR(table_no_outlier$Viscera_weight)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

table_no_outlier <- subset(table_no_outlier, table_no_outlier$Viscera_weight> Lower & table_no_outlier$Viscera_weight < Upper)
boxplot(table_no_outlier$Viscera_weight,
         ylab = "Viscera weight") 
```
```{r Shell_weight_outlier}
quartiles <- quantile(table_no_outlier$Shell_weight, probs=c(.25, .75), na.rm = FALSE) 
IQR <- IQR(table_no_outlier$Shell_weight)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

table_no_outlier <- subset(table_no_outlier, table_no_outlier$Shell_weight> Lower & table_no_outlier$Shell_weight < Upper)
boxplot(table_no_outlier$Shell_weight,
        ylab = "Shell weight")
```
```{r Rings_outlier}
quartiles <- quantile(table_no_outlier$Rings, probs=c(.25, .75), na.rm = FALSE) 
IQR <- IQR(table_no_outlier$Rings)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

table_no_outlier <- subset(table_no_outlier, table_no_outlier$Rings> Lower & table_no_outlier$Rings < Upper)
boxplot(table_no_outlier$Rings,
        ylab = "Rings")
```
Seems like got rid of emissions, but it's not certain. 
Now let's look at the correlations. 
*Correlations*
```{r correlation}
ggcorr(table_no_outlier)
```
Cool, a strongly positive correlation is observed almost everywhere.

## 3 The standard deviation of the Length variable for molluscs of different sexes.
```{r Lenght_sd}
tapply(X = table_no_outlier$Length, INDEX = table_no_outlier$Sex, FUN = mean)

tapply(X = table_no_outlier$Length, INDEX = table_no_outlier$Sex, FUN = sd)
```

## 4 The percentage of molluscs that have a value of the Height variable less than 0,165.
```{r pr}
x <- table_no_outlier[table_no_outlier$Height >= 0.165, ]
short <- dim(x)
all <- dim(table_no_outlier)
pr <- 100*short[1]/all[1]
pr
```



## 5 Values of the variable Length that are higher than 92% of all observations.

```{r len more 92}
Length <- c(table_no_outlier$Length)
q <- quantile(table_no_outlier$Length, probs= c(0.92, 0.92), na.rm = FALSE)
q
```
## 6 Z-transformation of Length

```{r z-score}
m <- mean(Length)
s <- sd(Length)
Lenght_z_scores <- (Length-m)/s

```

Perhaps there is a standard feature for it, but I could not find it.

## 7 Comparison between the diameter of clams with the number of rings 5 and 15.

```{r compare}
five_rings <- subset(table_no_outlier, table_no_outlier$Rings == 5)
summary(five_rings$Diameter)
fiveteen_rings <- subset(table_no_outlier, table_no_outlier$Rings == 15)
summary(fiveteen_rings$Diameter)
f_rings <- subset(table_no_outlier, table_no_outlier$Rings == 5 | table_no_outlier$Rings == 15)
f_rings$Rings <- as.character(f_rings$Rings)

```
```{r compare_plot}
ggplot(f_rings, aes(x = Rings, y = Diameter, color = Rings)) +
  geom_boxplot() +
  theme_classic()
```
```{r compare_plot2}
ggplot(f_rings, aes(x = Rings, y = Diameter, color = Rings)) + 
  geom_dotplot(binaxis='y', stackdir='center')
```
You can choose which one to use in your publication. 
Conclusions: Diameter of mollusks has a positive correlation with the number of rings. The median for molluscs with 5 rings is 0.2100, and with 15 rings 0.4600. The difference is more than twofold. 

## 8  Diameter and Whole_weight
```{r DW_plot}
cor.test(table_no_outlier$Diameter, table_no_outlier$Whole_weight)
```

There is a strong positive correlation.

```{r cor_plot}
plot(table_no_outlier$Diameter, table_no_outlier$Whole_weight)
Diameter <- table_no_outlier$Diameter
Whole_weight <- table_no_outlier$Whole_weight
plot(Diameter, Whole_weight, pch = 19, col = "lightblue")
```
## 9  My suggestions 
Correlation with sex
I was curious to see if there was a correlation with gender, so I decided to check it out. 


```{r sex_Length_plot}
ggplot(table_no_outlier, aes(x = Length, y = Sex, color = Sex)) +
  geom_boxplot() +
  theme_classic()
```

```{r sex_Diameter_plot}
ggplot(table_no_outlier, aes(x = Diameter, y = Sex, color = Sex)) +
  geom_boxplot() +
  theme_classic()
```

```{r sex_Whole_weight}
ggplot(table_no_outlier, aes(x = Whole_weight, y = Sex, color = Sex)) +
  geom_boxplot() +
  theme_classic()
```
```{r sex_Shucked_weight}
ggplot(table_no_outlier, aes(x = Shucked_weight, y = Sex, color = Sex)) +
  geom_boxplot() +
  theme_classic()
```
```{r sex_Viscera_weight}
ggplot(table_no_outlier, aes(x = Viscera_weight, y = Sex, color = Sex)) +
  geom_boxplot() +
  theme_classic()
```


```{r sex_Shell_weight}
ggplot(table_no_outlier, aes(x = Shell_weight, y = Sex, color = Sex)) +
  geom_boxplot() +
  theme_classic()
```



Conclusions: There is no correlation between females and males, but there is a correlation between sexually mature and nonsexually mature individuals.

